---
title: "R Notebook"
author: "Bibek Sapkota"
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---

# Vector Operators using R

task 1: Create two vector a and b
```{r}
a <- 1:10
b <- 1:10
```

```{r}
class(a)
class(b)
```

task 2: This is same as task-1
```{r}
c <- seq(1:10)
d <- as.integer(c(1, 2, 3, 4, 5, 6, 7, 8, 9,10))
```

```{r}
class(c)
class(d)
```

task 3:Adding two vector
```{r}
(a+b)
```

## Using () for printing the output

```{r}
y <- seq(1, 10, length.out= 5)
```

task 1: printing the expression
```{r}
(y <- seq(1, 10, length.out= 5))
```

## Vector Operation using R

task 1: Multiplying two vector
```{r}
a <- 1:10
b <- 5

(a*b)
```

```{r}
a <- 1:10
b <- 5
c <- c(2,3,4,5,6,7,8,9,10,11)

class(a)
class(b)
class(c)
```

## Blinding two column vector
```{r}
ac = cbind(a,c)
print(ac)
```

```{r}
acb <- ac * b
print(acb)
```

```{r}
class(acb)
```

## Calculating mean

```{r}
d <- apply(acb, MARGIN = 1, FUN =mean)
print(d)
```

```{r}
(acbd <- cbind(acb, d))
```

```{r}
class(acbd)
```

```{r}
summary(acbd)
```
## Vector Recycling

```{r}
a <- 1:10
b <- 1:5

(a+b)
```

```{r}
a <- 1:10
b <- 1:7

(a+b)
```

## Function in R
```{r}
best_practice <- c("Let", "the", "compiter", "do", "the", "work")

print_words <- function(sentence){
  print(sentence[1])
  print(sentence[2])
  print(sentence[3])
  print(sentence[4])
  print(sentence[5])
  print(sentence[6])
}

print_words(best_practice)
```

```{r}
print_words(best_practice[-6])
```

## Deleting element from vector
```{r}
best_practice[-6]
```

```{r}
best_practice <- c("Let", "the", "compiter", "do", "the", "work")

print_words <- function(sentence){
  for (word in sentence){
    print(word)
  }
}

print_words(best_practice)
```

```{r}
print_words(best_practice[-6])
```
## Apply, lapply, sapply, vapply

task 1: apply
```{r}
a <- 1:10
b <- 10:20

df <- data.frame(cbind (a, b) )
```

```{r}
apply(df, MARGIN = 1, FUN = mean)
```

task 2: lapply
```{r}
a <- 1:10

lapply(a, MARGIN=1, FUN= mean)
```

task 3:sapply
```{r}
sapply(a, MARGIN =1 , FUN= mean)
```

task4:vapply
```{r}
data <- list(l1 = c(1, 2, 3, 4),
             l2 = c(5, 6, 7, 8))

vapply(data, max, double(1) )
```

## Conditional Statements
```{r}
y <-10
if (y<20){
  x <- "Too Low"
}else{
  x <- "Too High"
}

print(x)
```


```{r}
temp <- 35

if (temp <= 0){
    "freezing"
}else if (temp <= 10){
  "cold"
}else if (temp <=20){
  "cool"
}else if (temp <=30){
  "warm"
}else{
  "hot"
}
```

# Exploring covid datasets

1. Create data frame with these two column vectors in R Studio x = 1:30 y = x^3
```{r}
x <- as.numeric(1:30)
y <- x^3

df <- data.frame(x,y)

str(df)
```

2. Create plot of x and y variables in R Studio and interpret it carefully
```{r}
plot(x,y)
```

3. Get appropriate correlation coefficient of this data in R Studio and interpret it carefully
```{r}
corr <- cor.test(x= df$x, y=df$y, method="spearman")
corr
```

4. Transform the plot to linear using appropriate mathematical function in R Studio
```{r}
plot(log(df$x), log(df$y))
```

6. Create a new column vector z defined in the slide 18 of session two slide deck in R Studio.
```{r}
z <- c(1, 1, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3,  4, 4, 4, 4, 4, 5, 5, 5, 6, 6, 7)
```

7. Create a histogram of z variable in Rstudio and interpret it carefully
```{r}
hist(z)
```

8.Get summary statistics of z variable in R Studio and interpret it carefully
```{r}
summary(z)
```

9. Get box-plot of z variable in R Studio and interpret the result carefully
```{r}
boxplot(z)
```

10. Import “covnep_252days.csv” data in R Studio and describe the variables in it.

```{r}
packages_to_install <- c("readr")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(readr)
```

```{r}
convnep_df <- read_csv("covnep_252days.csv", col_names = TRUE, col_types = cols(date = col_date(format = "%m/%d/%Y")))
str(convnep_df)
```

11. Create a chart with “totalCases” variable in y-axis and “date” variable in the x-axis in R Studio, describe the process leading to the creation of this chart.
```{r}
plot(convnep_df$date, convnep_df$totalCases, type='l')
```

12. Get summary statistics of “totalCases” variable in R Studio and interpret it carefully
```{r}
summary(convnep_df$totalCases)
```

13. Create histogram of “newCases” variable in R Studio and interpret it carefully.
```{r}
hist(convnep_df$newCases)
```

14. Get summary statistics of “newCases” variable in R Studio and interpret it carefully.
```{r}
summary(convnep_df$newCases)
```

15. Get “box and whisker” plot of “newCases” variable in R Studio and interpret it carefully.
```{r}
boxplot(convnep_df$newCases)
```

16. Import “SAQ8.sav” data in R Studio and get frequency distribution (number and percentage of the attributes) of q01, q03, q06 and q08 variables on R Studio and interpret them carefully.
```{r}
packages_to_install <- c("haven")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(haven)
```

```{r}
saq8_df <- read_sav("SAQ8.sav")

str(saq8_df)
```

```{r}
head(saq8_df, 10)
```



```{r}
packages_to_install <- c("plyr")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(plyr)
```

```{r}
# Define a function
col_list <- list("q01", "q03", "q06", "q08")

# Create a function frequency table
frequency_table <- function(df, col_list) {
  for(item in col_list) {
    new_df <- count(df, item)
    new_df$percentage <- round(new_df$freq / sum(new_df$freq) * 100, 2)
    new_df$cumulative_percentage <- cumsum(new_df$percentage)
    print(new_df)
  }
}

frequency_table(saq8_df, col_list)
```

17. Import “MR_drugs.xls” data in R Studio and replicate multiple response frequency distribution as shown in the slide 35 of the session 2 slide deck.
```{r}
packages_to_install <- c("readxl")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(readxl)
```

```{r}
drugs_df <- readxl::read_excel("MR_drugs.xls")

str(drugs_df)
```

```{r}
head(drugs_df, 10)
```

```{r}
mr_drugs_df <- data.frame(
  N = colSums(drugs_df[4:10]),
  Percent = round(colSums(drugs_df[4:10]) / sum(drugs_df[4:10]) * 100, 2),
  "Percent of Cases" = round(colSums(drugs_df[4:10]) / nrow(drugs_df[4:10]) * 100, 2)
)

print(mr_drugs_df)
```

```{r}
# summary table as in spss
# import library for table

packages_to_install <- c("gt")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(gt)

```

```{r}
# create table
mr_drugs_df %>%
  gt()%>%
  tab_header(
    title="$Income Frequencies"
  ) %>%
  tab_spanner(label="Responses", columns = c(N, Percent)) 

#summary_rows(
 #   columns = everything(),
  #  fns = list(Total = ~mean(.))
  #)
  
```

#Pipe Operators

1. Initialize a vector
```{r}
packages_to_install <- c("magrittr")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(magrittr)

x <-c (0.109, 0.359, 0.63, 0.996, 0.515, 0.142, 0.017, 0.829, 0.907)

# Compute the logarithm of `x`, 
# return suitably lagged and iterated differences 
# compute the exponential function and round the result
round(exp(diff(log(x))), 1)
```


2. Using pipe operator
```{r}
x %>%
  log() %>%
  diff() %>%
  exp() %>%
  round(1)
```

## Different pipe operators


```{r}
#Other pipe operators 1
x <- rnorm(100)

(x %<>% abs %>% sort)
```

```{r}
#Other pipe operators 2
rnorm(200) %>%
  matrix(ncol = 2) %T>%
  plot %>% 
  colSums
```

```{r}
#Other pipe operators 2
rnorm(200) %>%
  matrix(ncol = 2) %T>%
  plot %>% 
  colSums
```

```{r}
#The above code is a shortcut for this code:
rnorm(200) %>%
  matrix(ncol = 2) %T>%
  { plot(.); . } %>% 
  colSums
```

```{r}
#Other pipe operator 3
data.frame(z = rnorm(100)) %$% 
  ts.plot(z)
```

```{r}
#More examples:
#Load the package, install if require!
packages_to_install <- c("babynames")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(babynames)
library(dplyr)
```


```{r}
data("babynames")
```

```{r}
sum(select(filter(babynames,sex=="M",name=="Taylor"),n))
```

```{r}
# Do the same but now with `%>%`
babynames%>%filter(sex=="M",name=="Taylor")%>%
  select(n)%>%
  sum
```

```{r}
#Assigning new variable and using compound assignment pipe operator:
# Load in the Iris data
iris <- read.csv(url("http://archive.ics.uci.edu/ml/machine-learning-databases/iris/iris.data"), header = FALSE)
```


```{r}
# Add column names to the Iris data
names(iris) <- c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species")
```

```{r}
# Compute the square root of `iris$Sepal.Length` and assign it to the new variable
iris$Sepal.Length.SQRT <- 
  iris$Sepal.Length %>%
  sqrt()
```

```{r}
#Compound pipe operator:
# Compute the square root of `iris$Sepal.Length` and assign it to the same variable
iris$Sepal.Length %<>% sqrt
```

```{r}
#The tee operator:
set.seed(123)
rnorm(200) %>%
  matrix(ncol = 2) %T>%
  plot %>% 
  colSums
```

```{r}
#Exposing pipe operator: comes handy when "data" argument is not needed in a function
iris %>% 
  subset(Sepal.Length > mean(Sepal.Length)) %$%
  cor(Sepal.Length, Sepal.Width)
```

## dplyr Package in R

```{r}
#install.packages("hflights")
packages_to_install <- c("hflights")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(hflights)
#Without pipe operators:
grouped_flights <- group_by(hflights, Year, Month, DayofMonth)
flights_data <- select(grouped_flights, Year:DayofMonth, ArrDelay, DepDelay) 
summarized_flights <- summarise(flights_data, 
                                arr = mean(ArrDelay, na.rm = TRUE),     #Remove missing data!
                                dep = mean(DepDelay, na.rm = TRUE))     #Remove missing data!
```

```{r}
final_result <- filter(summarized_flights, arr > 30 | dep > 30) 
final_result
```

```{r}
# With pipe operators:
hflights %>% 
  group_by(Year, Month, DayofMonth) %>% 
  select(Year:DayofMonth, ArrDelay, DepDelay) %>% 
  summarise(arr = mean(ArrDelay, na.rm = TRUE), dep = mean(DepDelay, na.rm = TRUE)) %>% 
  filter(arr > 30 | dep > 30)
```

```{r}
#ARRNGE data with dplyr and pipe operators:
#Ascending order
iris %>%
  select(starts_with("Sepal")) %>%
  filter(Sepal.Length >=70) %>%
  arrange(Sepal.Length)     #Sort data in ascending order
```


```{r}
#Descending order:
iris %>%
  select(starts_with("Sepal")) %>%
  filter(Sepal.Length >=70) %>%
  arrange(desc(Sepal.Length))       #Sort data in descending order
```

```{r}
#MUTATE with dplyr and pipe operators:
iris %>%
  select(contains("Sepal")) %>%
  mutate(Sepal.Area = Sepal.Length * Sepal.Width)
```

```{r}
iris %>%
  select(ends_with("Length")) %>%
  mutate(Length.Diff = Sepal.Length - Petal.Length)
```

```{r}
iris %>%
  select(ends_with("Length"), Species) %>%
  rowwise() %>%
  mutate(Length.Diff = Sepal.Length - Petal.Length)
```

```{r}
iris %>%
  select(contains("Sepal"), Species) %>%
  transmute(Sepal.Area = Sepal.Length * Sepal.Width)

```

## We must use R markdown syntax R markdown with knitr and kable,

```{r}
knitr::kable(head(mtcars), digits = 2, align = c(rep("l", 4), rep("c", 4), rep("r", 4)))
```


```{r}
packages_to_install <- c("xtable")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(xtable)
```

```{r}
print(xtable(head(mtcars)), type = 'html')
```

# Packages in R

1. Install and import dplyr,rjson  package
```{r}
packages_to_install <- c("dplyr", "rjson")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(dplyr)
library(rjson)
```

2. Help on any install pacakage
```{r}

#help(rjson)
#or
#?rjson

```

3. Using the rjson library to read content from json file
```{r}
data <- fromJSON(file="jason_data.json") 
print(data) 
```

4. Convert to dataframe 
```{r}
jason_data_frame <- as.data.frame(data) 
print(jason_data_frame) 
```

5. 
```{r}
jason_data_frame$Salary <- as.numeric(jason_data_frame$Salary) 
summary(jason_data_frame$Salary) 
```

6. Plotting the histogram 
```{r}
hist(jason_data_frame$Salary) 
```

















