---
title: "Data Science Lab 2"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_notebook: default
---

# Variables 

task 1: Created variables x and y . Assign the product of x and y to z and Print z.
```{r}
x <- 2+5
y <- 6
z <- x*y
z
```

task 2: Created variables a and b and divide a/b then assign the result to c and Print c.
```{r}
a <- 10-7
b <- 4
c <- a/b
c
```
# Functions

task 1: Created the variable x and Assign the function (c),"c is a one dimensional array" to x, and print variable x.
```{r}
x = c(100,200,300)
x
```

# Data Types in R

task 1: Created the variable x and checked the class of x.
```{r}
x =5
class(x)
```

task 2:Checked the variable x is integer.
```{r}
is.integer(x)
```


# Vectors, Matrices, Arrays, Lists and DataFrames

task 1: Created the variable xx and Assign matrix with dimensions with 3rows and 2cols.
```{r}
xx = matrix(1:6, nrow=3, ncol =2)
xx
```

task 2: Print the class of variable xx.
```{r}
class(xx)
```

task 3:Checked the variable xx is vector or not.
```{r}
is.vector(xx)
```

task 4:checked the variable xx is matrix or not.
```{r}
is.matrix(xx)
```

task 5: Checked the length of variable xx.
```{r}
length(xx)
```
task 5:Checked the dimension of the variable xx.
```{r}
dim(xx)
```

# Datasets for apply family 

task 1:Imported the dataset of mtcars. Print the head of mtcars which will print 1st 6rows.
```{r}
data("mtcars")
head(mtcars)
```

# Apply function

task 1: The function max is applied to calculated row wise maximum values.
```{r}
apply (t(beaver1),1,max)
```

task 2: The function mean is applied for each column and mean is calculated.
```{r}
apply(mtcars,2,mean)
```

task 3: Calculate the remainder when each element of the columns in the mtcars dataset is divided by 10 and display the first 6 rows 
```{r}
head(apply(mtcars,2,function(x) x%%10))
```

# lapply function

task 1: Created the one-dimensional vector and assign to movies, then apply movies to lower function using lapply  and checked the structure of each elements and print it.
```{r}
movies <- c("SPYDERMAN","BATMAN","VERTIGO","CHINATOWN")
movies_lower <-lapply(movies, tolower)
str(movies_lower)
```

task 2: Converted list into vector using unlist and then Display it.
```{r}
movies_lower <-unlist(lapply(movies,tolower))
str(movies_lower)
```

# sapply function

task 1: Created a dataset cars that is assign to dt then calculated the minimum value of each column in the cars dataset using both lapply (resulting in a list) and sapply (resulting in a simplified vector), and then displays the list of minimum values.
```{r}
dt <- cars
lmn_cars <- lapply(dt, min)
smn_cars <- sapply(dt, min)
lmn_cars
```

task 2: Displays the list of minimum values that has used sapply .
```{r}
smn_cars
```

task 3: The code calculates the maximum value of each column in the cars dataset using both 'lapply' (resulting in a list) and 'sapply' (resulting in a simplified vector), and then displays the list of maximum values.
```{r}
lmxcars <- lapply(dt, max)
smxcars <- sapply(dt, max)
lmxcars
smxcars
```

task 4:The 'avg' that calculates the average of the minimum and maximum values in a vector(x) and applies this function to each column of the dataset using 'sapply', and then displays the resulting vector of averages.
```{r}
avg <- function(x) {
( min(x) + max(x) ) / 2}
fcars <- sapply(dt, avg)
fcars
```

# Saving and Loading Data 

task 1:Creates a data frame named 'EU' containing columns for European country names and their corresponding populations and Displaying it.
```{r}
EUCountryNames = c("United Kingdom", "Germany","France","Italy")
EUPopulation = c(63843856, 82562004,64982894,61142221)
EU = data.frame(EUCountryNames,EUPopulation)
EU
```

task 2: Store the data on disk in the native R format RData.
```{r}
save(EU,file = "EUInfo.RData")
```


task 3:It shows absolute filepath representing the current working directory of the R process
```{r}
getwd()
```
task 4:Removed the created data frame and then load the one we have from the disk.
```{r}
rm(EU)
load("EUInfo.RData")
EU
```

task 5:It code saves the dataframe EU to a CSV file named "EUInfo.csv".
```{r}
write.csv(EU,"EUInfo.csv")
```

task 6:It reads the CSV file "EUInfo.csv" into a new data frame named 'EU2' and then displays its contents.
```{r}
EU2 = read.csv("EUInfo.csv")
EU2
```

task 7:Installing the packages "xlsx"
```{r}
# install.packages("xlsx")
```

task 8:Load the package "xlsx"
```{r}
library(xlsx)
```

task 9:It saves the dataframe EU to an Excel file named "EUInfoNew.xlsx" with the sheet name "Sheet1" using the write.xlsx() function from the "xlsx" package.
```{r}
write.xlsx(EU,"EUInfoNew.xlsx",sheetName="Sheet1")
```

task 10:It reads the Excel file "EUInfoNew.xlsx" from the sheet named "Sheet1" into a new data frame named EU3 and then displays its contents.
```{r}
EU3 = read.xlsx("EUInfoNew.xlsx",sheetName="Sheet1")
EU3
```

# Tutorial - 2

task 1: Get random 10 values from normal distribution
```{r}
x <- rnorm(10)
x
```

task 2: Display value x
```{r}
print(x)
```

task 3:list objects from the session
```{r}
ls()
```

task 4:Remove Object from the session
```{r}
rm(x)
```

task 5:Get 10 values from 1 to 10 and Display it.
```{r}
a <- c(1:10)
print(a)
```

task 6:Get 10 values from 1 to 10 and Display it.
```{r}
b <-1:10
print(b)
```

task 7: Create vector using : operator
```{r}
x <- c(1, 5, 4, 9, 0)
typeof(x)

length(x)
x <- c(1, 5.4, TRUE, "hello")
x
typeof(x)
```

task 8:Create vector using seq() function. specify step size and length of the vector
```{r}
seq(1, 3, by=0.2) 

seq(1, 5, length.out=4) 
```

# Accessing an element of vector

task 1:Accessing 3rd element then accessing 2nd and 4th element and accessing all but not 1st element
```{r}
x <- 1:7

x[3]    

x[c(2, 4)]     

x[-1]          
# cannot mix positive and negative integers
# x[c(2, -4)]
```
## Using logical vector as index

task 1: filtering vectors based on conditions. condition will repeat if there is less conditions and more variable
```{r}
x <- 1:5
x[c(TRUE, FALSE, FALSE, TRUE)]

x[x < 0] 
x[x > 0]
```
## Using Character vector as index
```{r}
x <- c("first"=3, "second"=0, "third"=9)
names(x)
x["second"]
```
# Create 

task 1:Create a list of strings
```{r}
list_size <- list("small", "medium", "big")
print(list_size)
```

task 2:Create x and y vector
```{r}
x <- as.numeric(1:30)
y <- x ^ 3
```

task 3:Create a dataframe using x and y vector and then viewing a structure of dataframe.
```{r}
df <- data.frame(x, y)

str(df)
```

## Create plot of x and y variables in R Studio and interpret it carefully

task 1:create a plot using R base package
```{r}
plot(df$x, df$y)
```

```{r}
corr <- cor.test(x = df$x, y = df$y, method="spearman")
corr
```

task 2:Use log transformation to the x and y variables.
```{r}
plot(log(df$x), log(df$y))
```

```{r}
corr <- cor.test(x = log(df$x), y = log(df$y), method="pearson")
corr
```

# R Matrix

task 1: creating the matrix with 3rows and 3 columns
```{r}
matrix(1:9, nrow = 3, ncol = 3)
```

```{r}
matrix(1:9, nrow = 3)
```

task 2:fill matrix row-wise
```{r}
matrix(1:9, nrow=3, byrow=TRUE)
```

task 3:Create matrix and then renaming its columns in X,Y,Z and Rows in A,B,C
```{r}
x <- matrix(1:9, nrow = 3, dimnames = list(c("X","Y","Z"), c("A","B","C")))
x
```
task 4:access column names and rownames
```{r}
colnames(x)
rownames(x)
```

task 5:change column and row names
```{r}
colnames(x) <- c("C1","C2","C3")
rownames(x) <- c("R1","R2","R3")
x
```

task 6:'cbind' 1st plot column wise and 'rbind' plot row wies
```{r}
cbind(c(1,2,3),c(4,5,6))
rbind(c(1,2,3),c(4,5,6))
```

task 7:Use dim to create matrix
```{r}
x <- c(1,2,3,4,5,6)
dim(x) <- c(2,3)
x
```

```{r}
x <- c(1,2,3,4,5,6)
class(x)

dim(x) <- c(2,3)
x
class(x)
```

## Access Elements of Matrix

task 1:Create a matrix with dimension 3x3
```{r}
x <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9), nrow = 3, ncol = 3, byrow = TRUE)
x
```

task 2:select rows 1 & 2 and columns 2 & 3
```{r}
x[c(1,2),c(2,3)]
```

task 3:leaving column field blank will select entire columns
```{r}
x[c(3,2),] 
```

task 4:leaving row as well as column field blank will select entire matrix
```{r}
x[,]  
```

task 5:select all rows except first
```{r}
x[-1,]  
```

# Lists
```{r}
x <- list("a" = 2.5, "b" = TRUE, "c" = 1:3)
x
```

```{r}
str(x)
```

## Accessing the list components

task 1:Create a list and assign to x and Display it
```{r}
x <- list(name = "John", age = 19, speaks = c("English", "French"))
x
```

task 2:access elements by name
```{r}
x$name
x$age
x$speaks
```

task 3:access elements by integer index
```{r}
x[c(1, 2)]
x[-2]
```

task 4:access elements by logical index
```{r}
x[c(TRUE, FALSE, FALSE)]
```

task 5:access elements by character index
```{r}
x[c("age", "speaks")]
```

task 7:Create a list
```{r}
x <- list(name = "John", age = 19, speaks = c("English", "French"))
x
```

task 8:access element by name using single bracket []
```{r}
x["age"]
```

task 9:check the type of the result (single bracket returns a list)
```{r}
typeof(x["age"])
```

task 10:access element by name using double bracket [[]]
```{r}
x[["age"]]
```

task 11:check the type of the result (double bracket returns the content)
```{r}
typeof(x[["age"]])
```

## Add a component in list

task 1:Create a list and display it
```{r}
x <- list(name = "Clair", age = 19, speaks = c("English", "French"))
x
```

task 2:assign a new element to the list using double brackets [[]] and print the updated list
```{r}
x[["married"]] <- FALSE
x
```

## Delete a component in list

task 1:Create a list and display it
```{r}
x <- list(name = "Clair", age = 19, speaks = c("English", "French"))
```

task 2:remove an element from the list using double brackets [[]]
```{r}
x[["age"]] <- NULL
x
```

task 3:print the structure of the updated list
```{r}
str(x)
```

task 4:remove an element from the list using $ notation
```{r}
x$married <- NULL
x
```

task 5:print the structure of the updated list
```{r}
str(x)
```
