---
title: "Text Processing (Lab-12)"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_notebook: default
---

# String Manipulation:

## List of String Manipulation Functions 

task 1: Assigning value to virables and then printing its type.
```{r}
text <- "san francisco" 
typeof(text) 

num <- c("24","34","36") 
typeof(num) 
```

task 2:Assing value to var3 and printing it.
```{r}
var3 <- paste("Var1","Var2",sep = "-") 
var3 
```
task 3: Starting with 1 connect with ? and ! and assigning - sep in middle.
```{r}
paste(1:5,c("?","!"),sep = "-") 
```

task 4:Assiging value to variables then assiging sep in middle.
```{r}
text= "England" 
cat(text,"USA",sep = "-") 
```

task 5:Assiging sep in middle of month name from 1 Jan to May
```{r}
cat(month.name[1:5],sep = " ") 
```

task 6:Changing 1to10 num into string using tostring.
```{r}
toString (1:10) 
```

task 7:Importing the library
```{r}
library(stringr) 
```

task 8:Assign value to variable
```{r}
str <- "Los Angeles, officially the City of Los Angeles and often known by its 
initials L.A., is the second-most populous city in the United States (after New 
York City), the most populous city in California and the county seat of Los An 
geles County. Situated in Southern California, Los Angeles is known for its Medi 
terranean climate, ethnic diversity, sprawling metropolis, and as a major center 
of the American entertainment industry." 
strwrap(str)
```

task 9:count number of characters 
```{r}
nchar(str) 
str_length(str)
```

task 10:convert to lower 
```{r}
tolower(str)  
str_to_lower(str) 
```

task 11:Replace strings 
```{r}
chartr("and","for",x = str) #letters a,n,d get replaced by f,o,r 
str_replace_all(string = str, pattern = c("City"),replacement = "state") #this is case sensitive 
```

task 12:Extract parts of string 
```{r}
substr(x = str,start = 5,stop = 11) 
```

task 13:Get difference between two vectors 
```{r}
setdiff(c("monday","tuesday","wednesday"),c("monday","thursday","friday")) 
```

task 14:Check if strings are equal 
```{r}
setequal(c("monday","tuesday","wednesday"),c("monday","tuesday","wednesday")) 
setequal(c("monday","tuesday","thursday"),c("monday","tuesday","wednesday")) 
```

task 15:Abbreviate strings 
```{r}
abbreviate(c("monday","tuesday","wednesday"),minlength = 3) 
```

task 16:spliting strings 
```{r}
strsplit(x = c("ID-101","ID-102","ID-103","ID-104"),split = "-") 
str_split(string = c("ID-101","ID-102","ID-103","ID-104"),pattern = "-",simplify = T) 
```

task 17:find and replace first match 
```{r}
sub(pattern = "L",replacement = "B",x = str,ignore.case = T) 
```

task 18:find and replace all matches 
```{r}
gsub(pattern = "Los",replacement = "Bos",x = str,ignore.case = T) 
```
### Question 1:Write a command to extract the first 5 characters of above-given text (string). 
```{r}
dt <- c("Soloman", "abcdef", "snakjs")
first_5_characters <- substr(dt, start = 1, stop = 5)
print(first_5_characters)
```


## Metacharacters

task 1:
```{r}
dt <- c("20","20$") 
grep(pattern = "20\\$",x = dt,value = T) 
```

task 2:sub() function to make the replacements. 
```{r}
dt <- c("may?","money$","and&") 
gsub(pattern = "[\\?-\\$-\\&]",replacement = "",x = dt) 
```


task 3: double backslash in a string, you'll need to prefix it with another double backslash to get detected
```{r}
gsub(pattern = "\\\\",replacement = "-",x = "Barcelona\\Spain") 
```
### Question 2: Write a code to capture the first two elements in dt <- c("may?, "money$" but not "and&").
```{r}
dt <- c("may?", "money$", "and&")
result <- grep("^(may\\?|money\\$)$", dt, value = TRUE)
print(result)
```

##  Quantifiers 

task 1:Assign value to variable
```{r}
names <- c("anna","crissy","puerto","cristian","garcia","steven","alex","rudy")
```

task 2:The symbol .? is known as a non-greedy quantifier. Being non-greedy, for a particular pattern to be matched, 
it will stop at the first match.
```{r}
#doesn't matter if e is a match 
grep(pattern = "e*",x = names,value = T) 

#must match t one or more times 
grep(pattern = "t+",x = names,value = T) 

#must match n two times 
grep(pattern = "n{2}",x = names,value = T) 
```

### Question 3:Write a regex pattern to match ‘c’ one or more times. 
```{r}
# Use grep to find elements containing 'c' one or more times
matches <- grep("c+", names, value = TRUE)

# Print the matching names
print(matches)
```

## Sequence

task 1:matching a digit 
```{r}
str <- "I have been to Paris 20 times" 
 
gsub(pattern = "\\d+",replacement = "_",x = str)
regmatches(str,regexpr(pattern = "\\d+",text = str)) 
```

task 2:match a non-digit 
```{r}
 
gsub(pattern = "\\D+",replacement = "_",x = str)  
regmatches(str,regexpr(pattern = "\\D+",text = str))
```

task 3:match a space - returns positions  
```{r}
gregexpr(pattern = "\\s+",text = str) 
```

task 4:match a non space  
```{r}
gsub(pattern = "\\S+",replacement = "app",x = str)
```

task 5:match a word character  
```{r}
gsub(pattern = "\\w",replacement = "k",x = str) 
```

task 6:match a non-word character  
```{r}
gsub(pattern = "\\W",replacement = "k",x = str)
```

## Character Classes 

task 1:extract numbers 
```{r}
str <- "20 people got killed in the mob attack. 14 got severely injured" 

regmatches(x = str,gregexpr("[0-9]+",text = str)) 
```

task 2:extract without digits 
```{r}
regmatches(x = str,gregexpr("[^0-9]+",text = str)) 
```

## Lecture examples: 

task 1:Checks if strings match three characters followed by a period.
```{r}
str<- c("cat.", "896.", "?=+.", "abc1") 
grepl(pattern = "...\\.",x=str) 
```

task 2:Checks if strings have 'a's, followed by 'b's, ending with 'c's.
```{r}
str<- c("aaaabc", "aabbbc", "aacc", "a") 
grepl(pattern = "a+b*c+",x=str) 
```

task 3:Checks if strings have 'waz' followed by 3-4 'z's and 'up'.
```{r}
str<- c("wazzzzup", "wazzzup", "wazup") 
grepl(pattern = "waz{3,4}up",x=str) 
```

task 4:Selects strings with lowercase, optional period, space, uppercase, and digit.
```{r}
str<-c("A. B","c! d", "e f", "g. H3", "i? J", "k L") 
grep(pattern ="[a-z]\\.?\\s+[A-Z]\\d*",x=str) 
```


task 5:Selects strings in mm/dd/yyyy or mm/dd/yy format.
```{r}
str<- c("09/01/2016", "09/21/16", "12/25/2016", "12/05/16") 
grep(pattern ="^([0-2][0-9]|(3)[0-1])(\\/)((0)[0-9]|(1)[0-2])(\\/)\\d{4}$", x=str)
```

### Question 4:Write a code to match only the first digit in dt <- c("75 to 79", "80 to 84", "85 to 89").
```{r}
dt <- c("75 to 79", "80 to 84", "85 to 89")
first_digits <- gsub(pattern = "(\\d+).*", replacement = "\\1", x = dt)
print(first_digits)
```










