---
title: "Introduction to Data Science Lab"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_notebook: default
---

# A calculator

task 1: Add
```{r}
2+5
```

task 2: Subtract
```{r}
10-6
```

task 3:Multiply
```{r}
100*5
```

task 4:Divide
```{r}
1000/5
```

task 5: Square
```{r}
2^3
```

task 6: Combine operations
```{r}
(2+3)*5
2+3*5
2+(3*5)
```

# Variables

task 1: Assiging value to x and y then multiplying and assiging in to z and Displaying results.
```{r}
x = 2+5
y = 6
z = x*y
z
```

task 2: Doing the same thing as task1 but using <- instead of =.
```{r}
a <- 10-7
b <- 4
c<-a/b
c
```

# Functions

```{r}
x = c(100,200,300)
x
```

#  Data Types in R

task 1:Assigning 5 to x then finding it class and checking is it is.integer or not.
```{r}
x = 5
class(x)
is.integer(x)
```

task 2: changing class of x as integer and then checking as is.integer. 
```{r}
 x = as.integer(5)
 class(x)
 is.integer(x)
```

task 3: Assigning name to variable my.name and checking its class
```{r}
 my.name <-"Mohamed Gaber"
 class(my.name)
```

task 4: Storing  Logical variables that store TRUE or FALSE and Displaying it and checking it class.
```{r}
x = 5
 y = 7
 z = x != y
 z

 class(z)
```














