---
title: "R Diamonds Dataset"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

# Install and Import ggplot2, factoextra and psych packages

```{r}
packages_to_install <- c("ggplot2", "factoextra","psych")

for (package_name in packages_to_install) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
  }
}

library(ggplot2)
library(factoextra)
library(psych)
```

# Loading the data into R

```{r}
data("diamonds")
```

task 1: Displaying first 6 rows data

```{r}
head(diamonds)
```

task 2: Displaying last 6 rows of data

```{r}
tail(diamonds)
```

task 3: Displaying more than 6 rows of data

```{r}
head(diamonds,10)
```

task 4: Checking the structure of data

```{r}
str(diamonds)
```

task 5: Checking the dimension of data

```{r}
dim(diamonds)
```

task 6: Summarizing the data

```{r}
summary(diamonds)
```

task 7: Describing the data

```{r}
 describe(diamonds)
```

# Visualization of diamonds dataset

task 1:Building the histogram of carat

```{r}
hist(diamonds$carat)
```

task 2:Using color to color the bin

```{r}
hist(diamonds$carat, col='aquamarine', ylab = "")
```

task 3:Changing the label of x-axis

```{r}
 hist(diamonds$carat, col='aquamarine', xlab = "Carat", ylab= "No of diamons")
```

task 4: Creating a box plot of all variables in dataset

```{r}
boxplot(diamonds)
```

task 5: Creating the box plot of only single variable (say diamonds)

```{r}
boxplot(diamonds$price)
```

task 6:Filling in the label on x-axis and y-axis

```{r}
boxplot(diamonds$price, 
        xlab="price", 
        ylab="Values")
```

task 7:Displaying Boxplot

```{r}
 boxplot(carat ~ price , data = diamonds,
 xlab = "price",
 ylab = "Carat",
 main = "Boxplot of carat by price")
```

task 8:Plotting the Scatter plot

```{r}
 plot(diamonds$carat, diamonds$clarity)
```

task 9: Calculating the distribution of transmissions

```{r}
trans_dist <- table(diamonds$color)
```

task 10:Creating pie chart

```{r}
pie(trans_dist,
 labels = c("D","E","F", "G","H" , "I", "J"),
 main = "colors of diamonds")
```

task 11:Using Scatter plot matrix for Displaying selected variables

```{r}
pairs(diamonds[, c("carat", "price", "depth", "table")],
 main = "Scatterplot Matrix")
```

task 12:Ploting Histogram of carat

```{r}
ggplot(diamonds, aes(x = carat),) +
 geom_histogram()
```

task 13:Using color to color the bin

```{r}
ggplot(diamonds, aes(x = carat, fill = cut)) +
  geom_histogram(bins = 30, position = "identity", alpha = 0.5) +
  theme_minimal() +
  labs(title = "Distribution of Diamond Carat Sizes by Cut",
       x = "Carat",
       y = "Count",
       fill = "Cut")
```

task 14:Changing the label of x-axis

```{r}
ggplot(diamonds, aes(x = carat)) +
 geom_histogram(fill = 'aquamarine', bins = 5) +
 xlab("")
```

task 15:Createing the box plot of only single variable

```{r}
ggplot(diamonds, aes(x= factor('depth'),y=depth)) +
 geom_boxplot()+
 ylab("Value") +
 xlab("depth")
```

task 16:Box ploting of carat by price

```{r}
 ggplot(diamonds, aes(x= factor(cut),y=color))+
 geom_boxplot()+
 xlab("carat") +
 ylab("price") +
 ggtitle("Box plot of carat by price")
```

task 17: Ploting the Scatter Plot

```{r}
 ggplot(diamonds, aes(x = carat, y = depth)) +
 geom_point()
```

# Applying K-means Clustering

task 1:set random seed

```{r}
set.seed(7696)
```

task 2: checking the class and dimension of data

```{r}
class(diamonds)
dim(diamonds)
```

task 3:filter data by column name

```{r}
filtered_data1= diamonds[, c("price","carat")]
```

task 4:Visualizing the data

```{r}
ggplot(diamonds, aes(x= price, y= carat))+ 
  geom_point()
```

task 5:Finding 5 clusters using k-means clustering and viewing the details

```{r}
clusters=kmeans(filtered_data1, 5)
clusters
```

task 6: Visulizing the cluster

```{r}
cluster_index = as.factor(clusters$cluster)
ggplot(filtered_data1, aes(x=price, y=carat, color= cluster_index )) +
  geom_point()
```

```{r}
fviz_cluster(clusters, data= filtered_data1, geom= c("point"))
```

task 7:Filtering the data

```{r}
filtered_data2= diamonds[, c("depth","price","carat")]
```

task 9:Finding 5 clusters using k-means clustering and viewing the details

```{r}
cluster2= kmeans(filtered_data2, 5)
cluster2
```


```{r}
fviz_cluster(cluster2, data= filtered_data2, geom= c("point"))
```



task 11:Finding 5 clusters using k-means clustering and viewing the details

```{r}
cluster3= kmeans(filtered_data2, 5 , nstart=25)
cluster3
```

task 12:Visulizing the cluster

```{r}
fviz_cluster(cluster3, data= filtered_data2, geom= c("point"))
```
