---
title: "Data Wrangling (Lab-7)"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_notebook: default
---

# Data Manipulation (Part-1)

task 1:
```{r}
library(tidyverse)

demo_data <- read_csv("yrbss_demo.csv")
demo_data
```

## Filter Rows

### Filter by position
```{r}
slice(demo_data, 1:6)

demo_data %>% slice(1:6)
```

### Filter by logical criteria

task 1:Filter rows with age less than 15
```{r}
demo_data[demo_data$age<15,]

demo_data %>% filter(age<15)
```

task 2:Filter rows using equality
```{r}
demo_data %>% filter(sex == "Female")
demo_data %>% filter(!(grade == "9th"))
```

task 3: Filter rows using OR operator
```{r}
demo_data %>% filter((bmi<15) | (bmi>50))
```

task 4:Filter rows using %in% operator.
```{r}
demo_data %>% filter(grade %in% c("10th", "11th"))
```

task 5:Filter rows by multiple-column based criteria.
```{r}
demo_data %>% filter(bmi<20, stweight<50, sex == "Male")
```

#### Q1:Write the code to filter rows with percentage of bmi/stweight < 0.5.
```{r}
demo_data %>% filter((bmi / stweight) < 0.5)
```


### Filter within a selection of variables.

task 1:Creating a temp_data by only using columns bmi and stweight
```{r}
temp_data <- demo_data %>% select(bmi, stweight)
```

task 2:Filtering rows where all variables greater than 2.4
```{r}
temp_data %>% filter_all(all_vars(.> 2.4))
```

task 3:Filtering rows where any variables are greater than 2.4
```{r}
temp_data %>% filter_all(any_vars(.> 2.4))
```

task 4:Filters rows in demo_data where both race values are "White".
```{r}
demo_data %>% filter_at(vars(starts_with("race")), all_vars(. == "White"))
```

task 5: Filter rows in demo_data where any of race values equals "Asian"
```{r}
demo_data %>% filter_at(vars(starts_with("race")), any_vars(. == "Asian"))
```

task 6:Filtering rows Which have none values for all numeric variables.
```{r}
demo_data %>% filter_if(is.numeric, all_vars(!is.na(.)))
```

### Select random rows from a dataframe

```{r}
set.seed(1234)

demo_data %>% sample_n(5, replace = FALSE)

demo_data %>% sample_frac(0.05, replace = FALSE)
```

### Select top n rows ordered by a variable

task 1:Select the top 5 rows ordered by bmi
```{r}
demo_data %>% top_n(5, bmi)
```

task 2:Select the top5 rows ordered by bmi in all grades
```{r}
demo_data %>% group_by(grade) %>% top_n(5, bmi)
```

## Column operations

### Select columns 
```{r}
demo_data[, c("record","age","sex")]

demo_data %>% select(record, age, sex)
demo_data %>% select(record:sex)
```

#### Q2: Explain the meaning of the output of: demo_data %>% select(-contains("r"))? 



### Rename columns
```{r}
demo_data %>% rename(id = record)
```

### Add a new column
```{r}
demo_data %>% mutate(height_m = sqrt(stweight / bmi))
```
















