---
title: "Association Rule Mining"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_notebook: default
---

task 1: Install package
```{r}
#install.packages("arules")
library("arules")
```

task 2:Importing and loading the data.
```{r}
dataAR <- read.csv("weather.csv")
dataAR
```

task 3: Applying apriori 
```{r}
rules <- apriori(dataAR)
```

task 4: Inspect the dataset
```{r}
inspect(rules)
```

task 5: Using Apriori algorithm mines association rules in dataAR with minimum 30% support and 30% confidence.
```{r}
rules <- apriori(dataAR, parameter =list(supp=0.3, conf=0.3) )
```

task 6:Display detailed information of the mined association rules.
```{r}
inspect(rules)
```

task 7:Finds association rules in dataAR with 30% support, 30% confidence, and minimum length of 2.
```{r}
rules <- apriori(dataAR, parameter =list (supp=0.3, conf=0.3, minlen=2 ) )
```

task 8: installing and loading the package
```{r}
#install.packages("arulesViz")
library("arulesViz")
```

task 9:Ploting rules
```{r}
plot(rules)
```

task 10:Using alternative graph
```{r}
plot(rules, method="graph", control = list(type="items"))
```



