---
title: "PCA"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_notebook: default
---

# PCA

## Loading the dataset and printing its first 6th values
```{r}
df <- iris[, -5]    
head(df) 
```

## Center and scale the data 
```{r}
df.scaled <- scale(df, center = TRUE, scale = TRUE) 
```

## Compute the correlation matrix: 
```{r}
# 1. Correlation matrix 
res.cor <- cor(df.scaled) 
round(res.cor, 2) 
```

## Calculate the eigenvectors/eigenvalues of the correlation matrix: 
```{r}
# 2. Calculate eigenvectors/eigenvalues 
res.eig <- eigen(res.cor) 
res.eig 
```

##  Compute projection matrix 

task 1:Define the dimensions of new data set 
```{r}
k <- 3 
```

task 2:Compute projection matrix (W) 
```{r}
W <- res.eig$vectors[,1:k] 
```


##  compute the new dataset: 

task 1: Using matrix calculation
```{r}
df.new = df.scaled %*% W 
```

task 2:Assign new column names 
```{r}
colnames(df.new) <- c("PC1", "PC2", "PC3") 
head(df.new) 
```

## Alternative method to compute new dataset 

task 1: Transpose the intial data  
```{r}
df.scaled.t <- t(df.scaled) 
```

task 2: Transpose the projection matrix 
```{r}
W.t <- t(W)  
```

task 3: Compute transpose of new dataset
```{r}
df.new.t <- W.t %*% df.scaled.t 
```

task 4: Compute the new dataset 
```{r}
df.new <- t(df.new.t) 
```

task 5: Assign new column names 
```{r}
colnames(df.new) <- c("PC1", "PC2", "PC3") 
head(df.new) 
```









