---
title: "T-test"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_notebook: default
---

task 1: Load built-in sleep data set
```{r}
sleep
```

task 2:Making a wide version of the sleep data; below we’ll see how to work with data in both long and wide formats
```{r}
sleep_wide <- data.frame(
    ID=1:10,
    group1=sleep$extra[1:10],
    group2=sleep$extra[11:20]
)
sleep_wide
```

## Compare two samples using T-test

task 1:Welch t-test
```{r}
t.test(extra ~ group, sleep)
```

task 2: Student t-test
```{r}
t.test(extra ~ group, sleep, var.equal=TRUE)
```

### Q1:What is the p-value of the t-test? 
0.s- The p value of t-test is 0.07919

### Q2:If significance level (∝) equals 0.05, do you accept alternative hypothesis (or reject null hypothesis) 
Ans-  If the p-value is less than or equal to 0.05, you reject the null hypothesis and accept the alternative hypothesis, indicating statistical significance. If the p-value is greater than 0.05, you do not reject the null hypothesis, meaning there isn't enough evidence to support the alternative hypothesis.






