---
title: "EDA (lab-10 / Part-1)"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_notebook: default
---

# Exploratory data analysis 

## Importing csv files 
```{r}
MS_county_stops <- read.csv('MS_county_stops.csv') 
MS_traffic_stops <- read.csv('MS_trafficstops_bw_age.csv') 
```

## Summarizing the data
```{r}
str(MS_county_stops) 
```
```{r}
summary(MS_county_stops) 
```

## Plotting with ggplot2 

task 1: importing the library
```{r}
library(ggplot2)
library(tidyverse)
```

task 2: Ploting the data of ms_country_stops dataset.
```{r}
ggplot(data = MS_county_stops, aes(x = male, y = female)) + geom_point()
```
```{r}
MS_county_stops %>% ggplot(aes(x = male, y = female)) + geom_point()
```

task 3: Assign plot to a variable and drawing data with datapoints and ploting it using lines
```{r}
MS_plot <- ggplot(data = MS_county_stops, aes(x = male, y = female)) 
MS_plot + geom_point() 
MS_plot + geom_line() 
```

### Scatter plot 

task 1:
```{r}
ggplot(data = MS_county_stops, aes(x = male, y = female)) + geom_point(alpha= 0.3) 
```

task 2: Adding blue color to the plot
```{r}
ggplot(data = MS_county_stops, aes(x = male, y = female)) + geom_point(alpha= 0.3, color= "blue")
```

task 3:Adding another line graph in data plot.
```{r}
ggplot(data = MS_county_stops, aes(x = male, y = female)) + geom_point(alpha= 0.3, color= "blue")+ geom_abline() 
```

###  Boxplot

task 1:Ploting box plot of Ms_traffic_stops dataset
```{r}
boxplot(driver_age~violation, data = MS_traffic_stops) 
```

task 2:Rotating the x axis text
```{r}
boxplot(driver_age~violation, data = MS_traffic_stops, las = 2) 
```

task 3:Drawing boxplots using the ggplot function.  
```{r}
ggplot(MS_traffic_stops, aes(x = violation, y = driver_age)) + geom_boxplot()
```

task 4:Filtering the missing values of driver age and Drawing boxplot
```{r}
filtered_MS_traffic_stops <- MS_traffic_stops %>% filter(!is.na(driver_age)) 
ggplot(filtered_MS_traffic_stops, aes(x = violation, y = driver_age)) + geom_boxplot() 
```

task 5:Filtering Yazoo country and removing its null values
```{r}
Yazoo_stops <- MS_traffic_stops %>% filter(county_name == "Yazoo County", !is.na(driver_age))
```

task 6:Using ggplot we can draw boxplots with data points on it.
```{r}
ggplot(Yazoo_stops, aes(x = violation, y = driver_age)) + geom_boxplot() + geom_point() 
```

task 7:
```{r}
ggplot(Yazoo_stops, aes(x = violation, y = driver_age)) + geom_boxplot() + geom_jitter() 
```

task 8:Coloring the noise
```{r}
ggplot(Yazoo_stops, aes(x = violation, y = driver_age)) + geom_boxplot() + geom_jitter(alpha = 0.5, color = "tomato")
```

task 9:
```{r}
ggplot(data = Yazoo_stops, aes(x = violation, y = driver_age)) + geom_jitter(alpha = 0.5, color = "tomato") + geom_boxplot()
```

task 10:
```{r}
ggplot(data = Yazoo_stops, aes(x = violation, y = driver_age)) + geom_jitter(alpha = 0.5, color = "tomato") + geom_boxplot(alpha = 0) 
```

### Histograms 

task 1:Ploting a hist driagram of driver age
```{r}
hist(MS_traffic_stops$driver_age,col="gold")
```

task 2:Drawing a histogram using the “ggplot” function.  
```{r}
ggplot(MS_traffic_stops, aes(driver_age)) + geom_histogram()
```

task 3:
```{r}
ggplot(MS_traffic_stops, aes(driver_age)) + geom_histogram(col="red", fill="green", alpha=0.2)
```

task 4:Define bin count and Use break sequence with lower bound, upper bound and bin range 
```{r}
ggplot(MS_traffic_stops, aes(driver_age)) + geom_histogram(col="red", fill="green", alpha=0.2, bins=20) 
 
ggplot(MS_traffic_stops, aes(driver_age)) + geom_histogram(col="red", fill="green", alpha=0.2, breaks=seq(10, 50, by=2)) 
```

task 5:Adding titles and x and y label
```{r}
ggplot(MS_traffic_stops, aes(driver_age)) + geom_histogram(col="red", fill="green", alpha=0.2, breaks=seq(10, 50, by=2)) + labs(title="Histogram forAge", x="Age", y="Count") 
```

task 6: Following code segment count values from the y-axis low values should be in  green and that the higher values should appear in red. 
```{r}
ggplot(MS_traffic_stops, aes(driver_age)) + geom_histogram(alpha=0.2, breaks=seq(10, 50, by=2), col="red", aes(fill=..count..)) + scale_fill_gradient("Count", low="green", high="red") 
```

### Bar charts / Bar plots 

task 1:Display the frequency of each category and Draw bar chart
```{r}
table(MS_traffic_stops$violation) 

barplot(table(MS_traffic_stops$violation),col = "red", main="Number of violations") 
```

task 2:
```{r}
ggplot(MS_traffic_stops, aes(violation)) + geom_bar()
```

task 3:colouring the bars, we will use fill, instead of colour
```{r}
ggplot(MS_traffic_stops, aes(violation)) + geom_bar(fill = "green")
```

task 4:Mapping the values to different colours. 
```{r}
ggplot(MS_traffic_stops, aes(violation)) + geom_bar(aes(fill = driver_gender)) 
```

task 5:stretch the bars between 0 and 1, by setting the position parameter to ‘fill’. 
```{r}
ggplot(MS_traffic_stops, aes(violation)) + geom_bar(aes(fill = driver_gender), position = "fill") 
```

task 6:Adding another function “coord_flip()”. In some scenarios, flipping will make the plot more readable. 
```{r}
ggplot(MS_traffic_stops, aes(violation)) + geom_bar(aes(fill = driver_gender), position = "fill") + coord_flip() 
```


## Correlation Analysis 


```{r}
# Load movies data set 
movies <- read.csv('movies.csv') 
```

### Graphical Analysis 

```{r}
ggplot(data = movies, aes(x=revenue, y=budget)) + geom_point(alpha= 0.3, color=  "blue")+ geom_abline()
```

###  Quantitative Analysis 

```{r}
cor(movies$vote_average, movies$revenue) 
```

### Correlation Matrix 

task 1:Loading library
```{r}
#install.packages("GGally")

library(GGally) 
```

task 2L Displaying the heatmap
```{r}
ggcorr(movies, label= TRUE , label_alpha= 0.3) 
```

###  pairwise correlation 
```{r}
ggplot(data = movies, aes(x = revenue, y = vote_count)) + geom_point(alpha= 
0.3, color= "blue") 
```

```{r}
qplot(movies$revenue, movies$vote_count, data=movies , geom= c("point", 
"smooth" ), method= "lm", alpha= I (1/5), se= FALSE) 
```

