---
title: "Data Enrichment (Lab-10 / Part-2)"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_notebook: default
---

# Data Enrichment 
. 
## Deduplication 

task1 :Import the library 
```{r}
library(tidyverse) 
```


task 2:create a vector of numbers
```{r}
x <- c(1, 1, 4, 5, 4, 6) 
```

task 3: Checking the positions of duplicate elements
```{r}
duplicated(x) 
```

task 4:Extracting duplicate elements 
```{r}
x[duplicated(x)]
```

task 5: remove duplicated elements, use !duplicated(), where ! is a logical negation. 
```{r}

x[!duplicated(x)] 
```

task 6:Loading Inbulit iris dataset
```{r}
iris[!duplicated(iris$Sepal.Width), ] 
```

### Remove duplicate rows in a data frame 

task 1: Remove duplicate rows based on all columns:
```{r}
iris %>% distinct() 
```


task 2: Remove duplicated rows based on Sepal.Length  
```{r}
iris %>% distinct(Sepal.Length, .keep_all = TRUE) 
```

task 3:Remove duplicated rows based on Sepal.Length and Petal.Width 
```{r}
 iris %>% distinct(Sepal.Length, Petal.Width, .keep_all = TRUE) 
```

## Grouping


```{r}
# First, we have to define which column we are going to use to group. 
# In this case we use the species of the plant 
my_group <- group_by(iris, Species) 
# Now using “summarize_all”, we define which function we use to group the values. 
# For an example we can use mean.  
summarize_all(my_group, funs(mean)) 
```

task 2: Download and load the data set. 
```{r}
surveys <- read.csv("survey.csv")
surveys
```

task 3:
```{r}
surveys %>%  
group_by(sex) %>%  summarize(mean_weight = mean(wgt, na.rm = TRUE)) 
```

task 4: group by multiple columns
```{r}
surveys %>%  
group_by(sex, species) %>%  summarize(mean_weight = mean(wgt, na.rm = TRUE))
```


```{r}
surveys %>%  
group_by(sex, species) %>%  summarize(mean_weight = mean(wgt, na.rm = TRUE), 
min_weight = min(wgt, na.rm = TRUE)) 
```














